---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Reproducible Research: Peer Assessment 1
----------------------------------------

```{r libraries, echo=FALSE}
library(dplyr)
```

## Loading and preprocessing the data
- Create a folder for ProjectData
- Unzip file if required
```{r unzip-file}
path <- "ProjectData"
if (!file.exists(path)) {
        dir.create(path)
}
if (!file.exists(file.path(path, "activity.csv"))) {
        unzip("activity.zip", exdir=path)
}
data <- read.csv(file.path(path, "activity.csv"))
# Uses dplyr library
steps.by.day <- aggregate(steps ~ date, data=data, FUN=sum)
```

## What is mean total number of steps taken per day?
```{r histogram}
hist(steps.by.day$steps, breaks=10, xlab="Number of steps", main="Steps taken per day")
```

## What is the average daily activity pattern?
valid.steps <- filter(data, !is.na(steps))
grouped.steps <- group_by(valid.steps, interval)
interval.means <- summarise(grouped.steps, mean(steps))


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
